### 2018-02-08 chapters/01-intro.tex
add R code to synth

The literature on political and criminal violence has increased exponentially over the last decades. Although interstate wars have traditionally occupied a privileged position in this research agenda, scholars have broadened their scope to include a myriad of hitherto understudied phenomena into their research endeavours. Civil wars \citep[][]{collier2004greed,fearon2003ethnicity,kalyvas2006logic}, genocides \citep[][]{mamdani2014victims, power2013problem}, ethnic conflicts \citep[][]{kaufmann1996possible, montalvo2005ethnic, sambanis2001ethnic}, wartime sexual abuse \citep[][]{cohen2013explaining,wood2006variation,wood2009armed}, electoral violence \citep[][]{hoglund2009electoral,wilkinson2006votes}, state-sponsored killings \citep[][]{harff1988toward, krain1997state,krain2005international,uzonyi2014unpacking}, terrorism \citep[][]{de2005quality,bueno2007propaganda,pape2003strategic}, drug-related violence \citep[][]{holmes2006drugs,lessing2015logics,richani2013systems, shirk2010drug}, street gangs \citep[][]{franzese2016youth,jones2009youth,rodgers2006living,sobel1987direct}, and prison gangs \citep[][]{dias2011pulverizaccao,freire2014,skarbek2011governance,skarbek2012prison,skarbek2014social} have recently moved to the centre stage of the discipline. This dissertation contributes to this growing literature.

In order to clarify crucial aspects of my research topic, I employ an eclectic combination of research designs, ranging from statistical analysis to field experiments and qualitative case studies. This diversity of methods not only reflects the multiple aspects of organised violence, but it is also a pragmatic response to problems which are common in this area, such as incomplete data, biased information, and the practical impossibility of random assignment of treatment. By using an array of methodological tools, I hope to overcome some of these inherent difficulties.

Regarding the geographical scope of this dissertation, most of the following chapters deal with issues of violence in Latin America, specially in Brazil. According to the World Bank, Latin America is home to about 8\% of the global population, yet it accounts for more than 30\% of the world's homicides.\footnote{See: \url{https://goo.gl/d2WC3V}. Access: April 2017.} Moreover, the yearly ranking by the Citizen's Council for Public Security and Criminal Justice (Consejo Ciudadano para la Seguridad Pública y la Justicia Penal), a Mexican non-governmental organisation, shows that 43 of the 50 most violent cities in the world are located in Latin America, including all of the top 10.\footnote{For the complete ranking, see \url{http://www.seguridadjusticiaypaz.org.mx/biblioteca/prensa/summary/6-prensa/239-las-50-ciudades-mas-violentas-del-mundo-2016-metodologia}. Access: April 2017.} Given the severity of violence in the continent, Latin America was the natural candidate to be the primary focus of this study.

>>>>

The literature on political and criminal violence has increased exponentially over the last decades. Although interstate wars have traditionally occupied a privileged position in political science, scholars have broadened their scope to include a myriad of hitherto understudied phenomena into their research agendas. Civil wars \citep{collier2004greed,fearon2003ethnicity,kalyvas2006logic}, genocides \citep{mamdani2014victims, power2013problem}, ethnic conflicts \citep{kaufmann1996possible, montalvo2005ethnic, sambanis2001ethnic}, wartime sexual abuse \citep{cohen2013explaining,wood2006variation,wood2009armed}, electoral violence \citep{hoglund2009electoral,wilkinson2006votes}, state-sponsored killings \citep{harff1988toward, krain1997state,krain2005international,uzonyi2014unpacking}, terrorism \citep{de2005quality,bueno2007propaganda,pape2003strategic}, drug-related violence \citep{holmes2006drugs,lessing2015logics,richani2013systems, shirk2010drug}, street gangs \citep{franzese2016youth,jones2009youth,rodgers2006living,sobel1987direct}, and prison gangs \citep{dias2011pulverizaccao,freire2014,skarbek2011governance,skarbek2012prison,skarbek2014social} have recently moved from the margins to the centre stage of the discipline. The present dissertation contributes to this expanding field.

In order to clarify crucial aspects of my research topic, I employ an eclectic combination of research designs. The methods range from qualitative case studies to randomised experiments and machine learning algorithms. This diversity not only reflects the multiple aspects of organised violence, but it is also a pragmatic response to problems which are common in this area, such as incomplete data, reporting bias, and model uncertainty. By using an array of methodological tools, I hope to overcome some of these challenges.

Regarding the geographical scope of this dissertation, two of the following chapters deal with issues of violence in Latin America, specially in Brazil. According to the World Bank, Latin America is home to about 8\% of the global population, yet it accounts for more than 30\% of the world's homicides.\footnote{See: \url{https://goo.gl/d2WC3V}. Access: April 2017.} Moreover, the yearly ranking by the Citizen's Council for Public Security and Criminal Justice (Consejo Ciudadano para la Seguridad Pública y la Justicia Penal), a Mexican non-governmental organisation, shows that 43 of the 50 most violent cities in the world are located in Latin America, including all of the top 10.\footnote{For the complete ranking, see \url{http://www.seguridadjusticiaypaz.org.mx/biblioteca/prensa/summary/6-prensa/239-las-50-ciudades-mas-violentas-del-mundo-2016-metodologia}. Access: February 2018.} Given the severity of violence in the continent, Latin America was expected to be an important part of this study.
